---
title: "ADC_Data-Tidying_Practice"
author: "Antonia Androski"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries Chunk}
library(dplyr)
library(tidyr)
```
It matters what order you call packages in, since R does not require unique nomenclature for functions. To access 'filter' from the 'stats' package: 
'stats::filter()'

# Read in and clean up data

Read in salmon catch .csv:
```{r Reading in}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)

head(catch_original)
```
When R reads in .csv files, it creates a relational data model in a space-efficient way by assigning each observation a factor. The 'stringsAsFactors' argument makes sure that character objects' properties are conserved (not converted to numbers)

The pipe operator (%>%) takes the output of one statement and uses it as the input for the next statement. If the assignment operator (<-) reads as "gets," the pipe operator reads as "then."

## %>% %>% %>% %>% %>% %>% %>% Typin' and pipin'

Select the columns we want using 'select()'

```{r Cleaning}
catch_data <- catch_original %>% 
  #selecting by columns that you DO want: select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  #OR select by columns that you DON'T want: 
  select(-All, -notesRegCode)

head(catch_data)
```
```{r}
summary(catch_data)
```
For some reason, Chinook's class is character. Change the values in the Chinook column to numeric using 'mutate()'
```{r Mutating Chinook column}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```
This conversion produced NAs - investigate using 'which' and 'is.na'
```{r Investigate}
i <- which(is.na(catch_clean$Chinook))
i
```

```{r}
catch_original[i,]
#look at column 401, where there is an NA
```
The 401st row should be a 1, not I/l. Do not go back to the original data and correct it!!! Fix it with a script.

use 'mutate' again to change the I/l to a 1. Then use 'mutate' to coerce Chinook column to numeric.Don't just turn all NAs into 1 - this would affect the entire dataset.

```{r Fix and then Mutate Chinook column}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I" | Chinook == "l", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))
head(catch_clean)
```


# Tidy data

Move from wide format to tall format using 'pivot_longer'
```{r Tidying}
catch_long <- catch_clean %>% 
pivot_longer(cols = -c(Region, Year), 
  names_to = "species",
  values_to = "catch")

#-c indicates that we are not pivoting over region or year

head(catch_long)
```

```{r Undo the pivot}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species,
values_from = catch)

head(catch_wide)
```

'rename()' catch to thousands

```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch * 1000)   

  head(catch_long)

```


# Summarize data
Mean catch by region and species of salmon

```{r Calculate Mean for All species by Region}
mean_region <- catch_long %>% 
  group_by(Region, species) %>% 
  summarise(mean_catch = mean(catch),
    n_obs = n())

mean_region
```

Mean sockeye catch by region

```{r Calculate Mean by Species}
mean_sockeye_region <- catch_long %>% 
  filter(species == "Sockeye") %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))

mean_sockeye_region
```

Group by year, then display top 5 years of catches
```{r Group Mean by Year}
annual_catch <- catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(desc(total_catch))

annual_catch
```
# Joining Tables

```{r Read in Region Codes table}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)


head(region_defs)
```
Create the join (create a 3rd table instead of overwriting the left or right table)
```{r Create Join}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```
```{r Instead of including the by argument, you can just rename region to code so that they match}
region_defs <- region_defs %>% 
  rename(Region = code)

catch_joined <-  left_join(catch_long, region_defs, by = "Region")

head(catch_joined)
```

```{r Create a test data frame with a single column, then split (separate) into 3 columns for month-day-year}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, into = c("month", "day", "year"), "/", remove = FALSE) %>% 
#Then unite the month, day, and year in a different order into date_2
  unite(date_2, year, month, day, sep = "-")
```
There is a package that is made for manipulating/working with dates: lubridate

```{r}

```

