---
title: "Salmon Escapement"
author: "Antonia Androski"
date: "10/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)
```


```{r Read in Data}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", 
  method = "libcurl"), 
  stringsAsFactors = FALSE)

#The function tryCatch can be useful if you need to switch between working from the internet and working from local files - it downloads the file when you need it.
```

# Calculate Annual Escapement for Salmon Species

```{r, results = "hide"}
# Change sample dates to years but keep sample dates, change year to numeric, group columns and summarize by calculating total escapement and 5 most common species
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-", remove = FALSE) %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Year, SASAP.Region, Species) %>% 
  summarize(total_escapement = sum(DailyCount)) %>%
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Pink", "Coho"))
#The %in% operator does the same thing as: filter(Species == "Chinook" | Species == "Sockeye" | etc...)
#You can also create a species list vector before running this filter, e.g. species_list <- c("Chinook", "Sockeye", etc...). This might be good since you can may want the list for other purposes. 

head(annual_esc)
```

# Visualize escapement with ggplot

Total Escapement by Region and Species

```{r Plot, echo = FALSE}
ggplot(data = annual_esc, mapping = aes(x = Species, y = total_escapement, fill = SASAP.Region)) +
  geom_col()
#You would think that the first part of this would make a plot, but it just sets up the plot without any points. You have to specify the geometry to get points.
```

Kodiak Escapement Over Time by Species

```{r, echo = FALSE}
my_theme <- theme_bw() + theme(legend.position = "top")

ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
       mapping = aes(x = Year, y = total_escapement, color = Species)) +
geom_point() + geom_line() + scale_color_brewer(palette = "Dark2") + ylab("Escapement (no. fish)") + ggtitle("Kodiak Salmon Escapement") + my_theme
```

Escapement Over Time: All Regions

```{r, fig.width = 7, fig.height = 20, echo = FALSE}
fig1 <- ggplot(annual_esc, aes(x = Year, y = total_escapement, color = Species)) + geom_point() + geom_line() + facet_wrap(~SASAP.Region, scales = "free_y", ncol = 1) + my_theme

fig1
#saving the plot in a new folder, "figures"
ggsave("figures/escapement.png", fig1, width = 12, height = 7, units = "in")
```

# Map Sampling Locations

Create a Table with Unique Sampling Locations, Latitudes, and Longitudes

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

head(locations)
#distinct function pulls out unique values from rows/columns. drop_na is included in tidyr.
```

```{r}
datatable(locations)
```

```{r, leaflet map}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~ Longitude, lat = ~ Latitude, popup = ~ Location)
```

```{r A more complex leaflet map}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```

For static maps, the package SF is more useful than leaflet. ggplot is compatible with data that is generated with the SF package. ggmap can also be used to bring in tiles from the internet. See section 13.5 of the ADC training book. 